GravExplain: Continuous gravitational wave searches in a table-top experiment

### 2020-02-13 paper/paper_main.tex

% general gw intro
In 2015 gravitational waves were observed for the first time from the merger of two black holes in a binary system.~\cite{GW150914} 
The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe. 
Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers. 
In recent years there has been an increased public interest in gravitational-wave science. 
Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience.
Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance \han{add refs}.

% what are gws and how do we detect them...
Gravitational waves are a prediction of Einstein's theory of general relativity. 
They are disturbances in spacetime caused by the acceleration of massive objects. 
The effect of gravitational waves is to change distances; a ``stretch and squeeze'' effect. 
Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
The observatories themselves are extremely complex devices, however they are based on the Michelson interferometer. 
Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate this science to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO}\han{add exhibit info when available} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.

% continuous gws
To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
However, these observatories are also searching for continuous gravitational waves; persistent, periodic signals which should be present in the data for many months or years. 
A prime target for these searches are rotating neutron stars. 
So far many searches have been made of LIGO and Virgo data, however no detection has been made so far.

% overview
In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous gravitational waves to a general audience. 
The techniques used are very similar to those used to search for continuous waves in LIGO and Virgo data.
As an overview, Section~\ref{sec:ifo} details the interferometer, Section~\ref{sec:single_tone} describes observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech. We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.

>>>>

% general gw intro
In 2015 gravitational waves were observed for the first time from the merger of two black holes in a binary system.~\cite{GW150914} 
The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe. 
Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers. 
In recent years there has been an increased public interest in gravitational-wave science. 
Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience.
Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.

% what are gws and how do we detect them...
Gravitational waves are a prediction of Einstein's theory of general relativity. 
They are disturbances in spacetime caused by the acceleration of massive objects. 
The effect of gravitational waves is a change in distances; a ``stretch and squeeze'' effect. 
Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
Gravitational-wave observatories are extremely complex devices, however they are based on the Michelson interferometer. 
Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.

% continuous gws
To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
These observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencis. 
One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
LMXBs are binary systms containing a compact object (such as a neturon star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
It is a prime target for continuous wave searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.

% sound analogy
Here we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
The analogy of listening to the universe is often used to explain gravitational-wave observations; transient gravitational-wave signals have been converted into audio.
Continuous waves may emitt at constant frequencies, or they may wander slowly in frequency over time. 
This can be demonstrated with a table-top interferometer by playing audio of constant or slowly changing frequency tones to produce a simulation of a continuous wave signal. 
The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data, particularly for the wandering frequency demonstration~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.

In this work we detail our table-top intererometer design in Section~\ref{sec:ifo}, in Section~\ref{sec:single_tone} we describe observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech. We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
